Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-02 paper/experiments.tex
Experiments updated

Amazon.com, an online retail store, organizes its catalog of 
products in a mulitply-rooted hierarchy and provides textual product descriptions
for most products.  Products can be discovered by users
through query-based searching or through product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

@@

We evaluated our model, HSLDA, against the comparison models with a focus
on predictive performance. Predictive performance was measured with
standard metrics - sensitiviy (true positive rate) and 
1-specificity (false positive rate). We evaluate on the two aforementioned
datasets to demonstrate that our model generalizes to two different
domains.

To evaluate the performance of these models, we established a gold standard for comparison.
For each dataset, a held out set of 1000 documents and labels were
reserved for evaluation and predictive performance was evaluated against a standard derived
from the observed labeling. To make the comparison as fair as possible, 
ancestors of observed nodes were ignored, observed nodes 
were considered positive and descendents of observed nodes were considered to be negative. This method of 
defining positive and negative labels was chosen to be as fair as possible to all models
being compared.  In particular, since the sLDA model does not enforce the hierarchical
constraints, the models can be compared on a more equal footing by considering only on the observed labels
as being positive despite the fact that ancestors must also be positive. The gold standard 
defined in this way will likely lead to a slight overestimation of the number of
false positives. It is known that ICD-9 codes lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as false positives.
However, given that the label space is often large (as in our examples) it is a moderate assumption that
erroneous false positives should not skew results significantly.

Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
where $\hat{d}$ represents the test document. For efficiency,
the expectation of this probability distribution was estimated in the following way. Expectations
of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples from the posterior.
Using these expectations, we performed Gibbs sampling over the hierarchy to acquire predictive
samples for the documents in the test set.

The following are the predictive performance results for the clinical
data given a prior mean for the regression parameters of -1.6. 
The full HSLDA model had a true positive rate of 0.57 
and a false positive rate of 0.13, the sLDA model had a true positive
rate of 0.42 and a false positive rate of 0.07, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.39
and a false positive rate of 0.08.

@@

The following are the predictive performance results for the retail product
data given a prior mean for the regression parameters of -2.2. 
The full HSLDA model had a true positive rate of 0.85 
and a false positive rate of 0.30, the sLDA model had a true positive
rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.77
and a false positive rate of 0.16. These results follow a similar pattern to the clinical data.

To further explore this tradeoff between the true positive rate and the
false positive rate we evaluated predictive performance for a range of values
for two different parameters - the prior mean for the regression coefficients
and the threshold for the auxiliary variables.  The goal in this analysis
was to evaluate the performance of these models subject to more or less
stringent requirements for predicting positive labels. These two parameters 
have important related functions in the model. The prior mean in combination 
with the auxiliary variable threshold together encode the strength of the prior
belief that unobserved labels are likely to be negative. Effectively, the
prior mean applies negative pressure to the predictions and the auxiliary
variable threshold determines the cutoff.

For each model type, separate models were fit for each value of the 
prior mean of the regression coefficients.  In contrast, to evaluate
predictive performance as a function of the auxiliary variable threshold,
a single model was fit for each model type and prediction was evaluated
based on predictive samples drawn subject to different auxiliary variable
thresholds. These methods are significantly different since the prior mean
is varied prior to inference and the auxiliary variable threshold is varied
following inference.  However, as intended, they both highlight model performance
under more or less stringent requirements for predicting positive labels.

>>>>

Amazon.com, an online retail store, organizes its catalog of 
products in a mulitply-rooted hierarchy and provides textual product descriptions
for most products.  Products can be discovered by users
through free-text search and product category exploration. Top-level
product categories are displayed on the front page of the website and lower
level categories can be discovered by choosing one of the top-level categories.
Products can exist in multiple locations in the hierarchy.

@@

We evaluated our model, HSLDA, against the comparison models with a focus
on predictive performance on held-out data. Prediction performance was measured with
standard metrics - sensitiviy (true positive rate) and 
1-specificity (false positive rate).

To evaluate the performance of these models, we established a gold standard for comparison.
For each dataset, a held out set of 1000 documents and labels were
reserved for evaluation and predictive performance was evaluated against a standard derived
from the observed labeling. To make the comparison as fair as possible, 
ancestors of observed nodes were ignored, observed nodes 
were considered positive and descendents of observed nodes were considered to be negative. This method of 
defining positive and negative labels was chosen to be as fair as possible to all models
being compared.  In particular, since the sLDA model does not enforce the hierarchical
constraints, its performance was found to be so poor on joint prediction that it was not included.
The models can be compared on a more equal footing by considering only on the observed labels
as being positive despite the fact that ancestors must also be positive. The gold standard 
defined in this way will likely inflate the number of
false positives because the labels applied to any particular document are usually not as complete as they could be. 
ICD-9 codes lack sensitivity and their use as a
gold standard could lead to correctly positive predictions being labeled as false positives\citep{Birmetal2005}.
However, given that the label space is often large (as in our examples) it is a moderate assumption that
erroneous false positives should not skew results significantly.

Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
for each test document, $\hat{d}$. For efficiency,
the expectation of this probability distribution was estimated in the following way. Expectations
of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples from the posterior.
Using these expectations, we performed Gibbs sampling over the hierarchy to acquire predictive
samples for the documents in the test set. The true positive rate was calculated as the
 average expected labeling for gold standard positive labels. The false positive rate was calculated
as the average expected labeling for gold standard negative labels.

The following are the performance results for the clinical
data given a prior mean for the regression parameters of $\mu=-1.6$. 
The full HSLDA model had a true positive rate of 0.57 
and a false positive rate of 0.13, the sLDA model had a true positive
rate of 0.42 and a false positive rate of 0.07, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.39
and a false positive rate of 0.08.

@@

The following are the predictive performance results for the retail product
data given a prior mean for the regression parameters of $\mu=-2.2$. 
The full HSLDA model had a true positive rate of 0.85 
and a false positive rate of 0.30, the sLDA model had a true positive
rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
LDA and the regressions were fit separately had a true positive rate of 0.77
and a false positive rate of 0.16. These results follow a similar pattern to the clinical data.

As sensitivity and specificity can always be traded off we examined sensitivity for a range of values
for two different parameters - the prior means for the regression coefficients
and the threshold for the auxiliary variables.  The goal in this analysis
was to evaluate the performance of these models subject to more or less
stringent requirements for predicting positive labels. These two parameters 
have important related functions in the model. The prior mean in combination 
with the auxiliary variable threshold together encode the strength of the prior
belief that unobserved labels are likely to be negative. Effectively, the
prior mean applies negative pressure to the predictions and the auxiliary
variable threshold determines the cutoff.

For each model type, separate models were fit for each value of the 
prior mean of the regression coefficients.  This is a proper Bayesian
sensitivity analysis. In contrast, to evaluate
predictive performance as a function of the auxiliary variable threshold,
a single model was fit for each model type and prediction was evaluated
based on predictive samples drawn subject to different auxiliary variable
thresholds. These methods are significantly different since the prior mean
is varied prior to inference and the auxiliary variable threshold is varied
following inference.  However, as intended, they both highlight model performance
under more or less stringent requirements for predicting positive labels.
